---
title: "Exponential Distribution Simulation"
author: "Andre Treffeisen"
date: "Saturday, March 21, 2015"
output: pdf_document
---

## Exponential Distribution Simulation
Project Statistical Inference - Part 1
March 2015

The exponential distribution can be simulated in R with `rexp(n, lambda)` where
`lambda` $\lambda$ is the rate parameter. The mean of exponential distribution is 
$1/\lambda$ and the standard deviation is also $1/\lambda$. For this simulation,
we set $\lambda=0.2$. In this simulation, we investigate the distribution of
averages of 40 numbers sampled from exponential distribution with $\lambda=0.2$.

####Simulations
```{r, fig.height=5, fig.width=5}
set.seed(666)
lambda = 0.2
samplesize = 40
sims = c(1:1000)
means <- data.frame(x = sapply(sims, function(x) {mean(rexp(samplesize, lambda))}))
head(means)
```


####Sample Mean versus Theoretical Mean
The actual and theoretical mean for the sample data is
```{r}
mean(means$x)           # actual
1/lambda                # theoretical
```
which are quite close. This distribution is centered around the actual figure above. Please also see the diagram below which shows the theoretical mean highlighted with a vertical red line.

####Sample Variance versus Theoretical Variance
The actual and theoretical standard deviation is
```{r}
sd(means$x)             # actual
(1/lambda)/sqrt(40)     # theoretical
```
which are quite close.

The actual and theoretical variance for the sample data is
```{r}
var(means$x)             # actual
((1/lambda)/sqrt(40))^2  # theoretical
```
which are also quite close. 

####Distribution is normal
```{r}
# histogram of averages
hist(means$x, breaks=50, prob=TRUE,
     main="Distribution of averages of samples,
     drawn from exponential distribution with lambda=0.2",
     xlab="")
# averages of samples: denstity
lines(density(means$x))
# theoretical distribution center (vertical red line)
abline(v=1/lambda, col="red")
# averages of samples (red dotted line): theoretical density
x <- seq(min(means$x), max(means$x), length=100)
y <- dnorm(x, mean=1/lambda, sd=(1/lambda/sqrt(samplesize)))
lines(x, y, pch=22, col="red", lty=2)
# add description
legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("black", "red"))
```


```{r}
```

```{r}
```

```{r}
```

```{r}
```

