---
title: "Exponential Distribution"
author: "Andre Treffeisen"
date: "Saturday, March 21, 2015"
output: pdf_document
---

# Simulation Of Exponential Distribution

Project Statistical Inference - Part 1

March 2015

## Overview

The exponential distribution can be simulated in R with `rexp(n, lambda)` where
`lambda` $\lambda$ is the rate parameter. The mean of exponential distribution is 
$1/\lambda$ and the standard deviation is also $1/\lambda$. For this simulation,
we set $\lambda=0.2$. In this simulation, we investigate the distribution of
averages of 40 numbers sampled from exponential distribution with $\lambda=0.2$.

## Simulations
```{r, fig.height=5, fig.width=5}
set.seed(666)
lambda = 0.2
samplesize = 40
sims = c(1:1000)
means <- data.frame(x = sapply(sims, function(x) {mean(rexp(samplesize, lambda))}))
head(means)
```


### Sample Mean versus Theoretical Mean
The actual and theoretical mean for the sample data is
```{r}
mean(means$x)           # actual
tm <- 1/lambda          # theoretical
tm
```
which are quite close. This distribution is centered around the actual figure above. Please also see the diagram below which shows the theoretical mean highlighted with a vertical red line.

### Sample Variance versus Theoretical Variance
The actual and theoretical standard deviation is
```{r}
sd(means$x)                   # actual
tv <- (1/lambda)/sqrt(40)     # theoretical
tv
```
which are quite close.

The actual and theoretical variance for the sample data is
```{r}
var(means$x)             # actual
((1/lambda)/sqrt(40))^2  # theoretical
```
which are also quite close. 

### Distribution is normal
```{r}
# histogram of averages
hist(means$x, breaks=50, prob=TRUE,
     main="Distribution of averages of samples,
     drawn from exponential distribution with lambda=0.2",
     xlab="")
# averages of samples: denstity
lines(density(means$x))
# theoretical distribution center (vertical red line)
abline(v=1/lambda, col="red")
# averages of samples (red dotted line): theoretical density
x <- seq(min(means$x), max(means$x), length=100)
y <- dnorm(x, mean=1/lambda, sd=(1/lambda/sqrt(samplesize)))
lines(x, y, pch=22, col="red", lty=2)
# add description
legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("black", "red"))
```
The figure above shows that the histogram can be adequately aproximated with the normal distribution.
We have shown that the variance and mean of sample data close resembles those of normal distribution. Let's also try to match the confidence intervals which are calculated below:

```{r}
sm <- means$x
aci <- round(mean(sm) + c(-1,1) * 1.96 * sd(sm) / sqrt(samplesize), 3)
tci <- tm + c(-1, 1) * 1.96 * sqrt(tv) / sqrt(samplesize)
aci        # actual confidence interval
```


```{r}
tci
```
Both confidence intervals are also close.

```{r}
qqnorm(sm)
qqline(sm)
```

With the theoretical quantiles also matching closely the actual quantiles, it shows that the distribution is approximately normal.

